mruby Bytecode Loader Using Bluetooth in Multi-VM Environment

### 2016-02-09 EMSOFT2016/thesis.tex
update

{\bf Contributions}: This paper proposes mruby bytecode loader using Bluetooth and user-friendly multitasking.
The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item To improve the  work efficiency in software development on mruby on TECS: \mbox{}\\
\ mruby bytecode loader using Bluetooth helps developers develop software.
Developers do not need to insert and pull out an SD card repeatedly and also to restart an OS.
Thus, the software development efficiency is improved.
\item To effectively execute multiple mruby programs in concurrent or/and parallel: \mbox{}\\
\ There is a problem in multitasking of the current mruby on TECS.
It is that developers need to be well acquainted the RTOS if multiple mruby programs execute in multitasking.
In the proposed framework, it is easy for developers to use multitasking because multiple tasks are switched cyclically.
Synchronization of tasks is implemented.
Multitasked programing makes the range of software expand.
\end{enumerate}

{\bf Organization}: The paper is organized as follows.
Section \ref{sec:Background} introduces the basic technologies i.e. mruby, TECS and mruby on TECS.
Section \ref{sec:mruby bytecode loader using Bluetooth} and Section \ref{sec:Multitask} describe the design and implementation of the proposed framework in detail.
Section \ref{sec:Evaluation} evaluates the proposed framework, Section \ref{sec:Related work} discusses related work, and then Section \ref{sec:Conclusion} concludes this paper.

@@

\section{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.pdf}
    \caption{Detailed System Model of the proposed framework}
    \label{fig:system_model}
\end{figure}
This section describes an additional feature of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, all binary data including bytecodes are in an SD card.
Developers must insert and pull out the SD card every time the source programs are modified.
The OS in the target device should be also restarted.
It causes a low work efficiency that such a work must be repeated.
mruby bytecode loader using Bluetooth makes the developers' burden decrease.
The work that an SD card is in and out needs to be done just once in the beginning.

@@

\subsection{Component of mruby bytecode loader using Bluetooth}
The proposed framework provides components of mruby bytecode loader using Bluetooth to use the function.
A component of mruby bytecode loader using Bluetooth is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages RiteVM configuration such as automatically generates the bytecode in the build description.
This generated bytecode is prepared beforehand in the SD card such as mruby libraries, and different from a bytecode transferred with Bluetooth.

@@

\end{lstlisting}
\caption{Main code for mruby bytecode loader using Bluetooth}
\label{maincode_mrubybluetooth}
\end{figure}
\section{Multitask}
\label{sec:Multitask}
This section describes implementation of multitasking in the proposed framework.
mruby on TECS has supported multitasking.
However, multitask processing in mruby on TECS requires the knowledge of the RTOS (TOPPERS/HRP2) for developers.

@@

In the proposed framework, as an approach for multitask processing, {\it rotateReadyQueue}, a service call of $\mu$ITRON, is implemented.
{\it rotateReadyQueue} rotates the precedence of the task with the priority of the argument.
A cyclic handler calls {\it rotateReadyQueue}, which enables developers to develop applications without concern for multitasking.

\subsection{rotateReadyQueue}
The case is assumed that there are two tasks with the same priority, and both tasks are in an infinite loop.
In the current system, when one task is executed first, another task would not be executed.
That is because the task with first execution runs in the loop.

@@

\subsection{Component of Cyclic Handler}
Figure \ref{fig:cyclic_handler} shows a component diagram of the cyclic handler.
The components of cyclic handler consist of two components: CyclicHandler and CyclicMain.
CyclicHandler {\it cell} configures the cyclic handler based on $\mu$ITRON.
Cyclic handlers based on $\mu$ITRON are described in detail \cite{par:microITRON}.
The cyclic handler has five arguments: ID, attribute, cyclic time, cyclic phase and access pattern.
CyclicHandler {\it cell} has these arguments as attributes of the {\it cell}.
CyclicMain {\it cell} is a component to perform the processing body of a cyclic handler.
{\it rotateReadyQueue} is implemented as the body.
Figure \ref{celltype_cyclic_handler} shows tCyclicMain {\it celltype}, which has a {\it call} port, an {\it entry} port and an attribute.
The {\it call} port is connected with the {\it entry} port of the Kernel {\it cell} ({\it tkernel.eiKernel}) to call functions of the kernel. 
The attribute is used as an arguments of {\it rotateReadyQueue}.

Figure \ref{build_cyclic_handler} shows a build description that corresponds to the component diagram shown in Figure \ref{fig:cyclic_handler}.
In the part of CyclicHandler {\it cell}, configurations of a cyclic handler is described such as attribute, cyclicTime and cyclicPhase.
In this case, the cyclic handler is executed when it is generated because the attribute is {\it TA\_STA} that represents the cyclic handler is in an operational state after the creation.
The cyclic handler is executed every one msec.
In another part, the priority of CyclicMain {\it cell} is described.
RITEVM\_PRIORITY defines the priority of mruby tasks.
In the main of CyclicMain, {\it rotateReadyQueue} is implemented and the priority is passed as the argument.
If developers do not need the cyclic handler, the .cdl files shown as Figures \ref{celltype_cyclic_handler} and \ref{build_cyclic_handler} should be commented out such as {\it //import($<$tCyclicHandler.cdl$>$);}.
This ease comes from component-based development with TECS.

@@

\subsection{Synchronization of Multiple RiteVM Tasks}
In the proposed framework, RiteVMs read mruby bytecodes, and then execute the applications.
Eventflag, one of synchronous processing, is applied to synchronize the starting of multiple mruby applications.
Each task sets the flag pattern such as 0x01 (01) and 0x02 (10), and then waits the flag pattern, 0x3 (11), with AND.
This process can also apply to the case of the more tasks.
For example, in the case of the four RiteVM tasks, each task sets the flag pattern such as 0x01 (0001), 0x02 (0010), 0x04 (0100)  and 0x08 (1000), and then waits 0x0f (1111) with AND.

In the proposed framework, the set pattern and wait pattern are defined as attributes of the component as shown in Figure \ref{fig:Eventflag}.
This design such as {\it cEventflag\_set(ATTR\_setPattern)} enables the program without ``if" statements and reuses the identical .c file.
Developers do not need to modify the .c file because the .cdl files in accordance with the number of RiteVMs are prepared. 
In additon, the components of Eventflag are built with {\it [optional]} in TECS.
{\it [optional]} means that the codes are run only when the call port is connected.
The .c file does not be rewritten even if developers do not use Eventflag. 
These features are also advantages of component-based development.

@@

\section{Experimental Evaluation}
\label{sec:Evaluation}
This section mentions experimental results and their consideration.
To analyze the advantages of the proposed framework, the evaluations are performed as follows.
\begin{itemize}
    \item the comparison of the size and transfer time between an mruby application including mruby library and not
    \item the comparison of the application execution time with singletasking, co-routine, and multitasking
    \item the comparison of the overhead for each cyclic period of calling {\it rotateReadyQueue}
\end{itemize}

@@

\begin{table*}[t]
    \centering
    \caption{Comparion of the proposed and previous work}
    \begin{tabular}{c||c|ccccccc}
        & Bluetooth Loader & \shortstack{Call\\C function} & \shortstack{Legacy code of\\embedded system} & VM managenment & VM Scheduler & \shortstack{Synchronization of\\application} & Co-routine \\ \hline
        python-on-a-chip \cite{url:python-on-a-chip} &            &            &            &            &            &            & \checkmark \\
        Owl system \cite{par:owl}                    &            & \checkmark & Partially  &            &            &            & \checkmark \\
        eLua \cite{url:eLua}                         &            & \checkmark & Partially  &            &            &            & \checkmark \\
        mruby \cite{par:mruby}                       &            & \checkmark &            &            &            &            & \checkmark \\
        mruby on TECS \cite{par:mrubyonTECS}         &            & \checkmark & \checkmark & \checkmark &            &            & \checkmark \\
        Proposed framework                           & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
    \end{tabular}
    \label{tab:comparion}
\end{table*}
The open-source run-time systems for scripting languages have been proposed such as follow:
python-on-a-chip \cite{url:python-on-a-chip}, the Owl system \cite{par:owl}, eLua \cite{url:eLua}, mruby \cite{par:mruby}, \cite{url:mruby}, and mruby on TECS \cite{par:mrubyonTECS}.

@@

Table \ref{tab:comparion} shows a comparison between the proposed framework and previous work.

@@

The proposed framework is developed in component-base by TECS.
The facilities such as the loader, the cyclic handler, and the Eventflag are implemented as components.
Developers can easily add or remove the facilities as necessary.
Component-based development can increase the productivity and decrease the complexity.

>>>>

{\bf Contributions}: This paper proposes mruby bytecode loader using Bluetooth and user-friendly multitasking.
The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item To improve the  work efficiency in software development on mruby on TECS: \mbox{}\\
\ mruby bytecode loader using Bluetooth helps developers develop software.
Developers do not need to insert and pull out an SD card repeatedly and also to restart an OS.
 Thus, the software development efficiency is improved.
\item To effectively execute multiple mruby programs in concurrent or/and parallel: \mbox{}\\
\ There is a problem in multitasking of the current mruby on TECS.
It is that developers need to be well acquainted the RTOS if multiple mruby programs execute in multitasking.
In the proposed framework, it is easy for developers to use multitasking because multiple tasks are switched cyclically.
Synchronization of tasks is implemented.
Multitasked programing makes the range of software expand.
\item To focus on the advantage of component-based developments: \mbox{}\\
\ The proposed framework has been developed with component-based.
Component-based development gives various benefits.
This paper shows the specific examples for the benefits of component-based development. 
\end{enumerate}

{\bf Organization}: The paper is organized as follows.
Section \ref{sec:Background} introduces the basic technologies i.e. mruby, TECS and mruby on TECS.
Section \ref{sec:Design and Implementation} describe the design and implementation of the proposed framework in detail.
Section \ref{sec:Evaluation} evaluates the proposed framework, Section \ref{sec:Related work} discusses related work, and then Section \ref{sec:Conclusion} concludes this paper.

@@

\section{Design and Implementation}
\label{sec:Design and Implementation}
Figure \ref{fig:system_model} shows the detailed system model of the proposed framework.
Each mruby application bytecode transferred from the host is received by the loader in the RiteVM.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.pdf}
    \caption{Detailed System Model of the proposed framework}
    \label{fig:system_model}
\end{figure}
\subsection{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
This section describes an additional feature of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, all binary data including bytecodes are in an SD card.
Developers must insert and pull out the SD card every time the source programs are modified.
The OS in the target device should be also restarted.
It causes a low work efficiency that such a work must be repeated.
mruby bytecode loader using Bluetooth makes the developers' burden decrease.
The work that an SD card is in and out needs to be done just once in the beginning.

@@

\subsubsection{Component of mruby bytecode loader using Bluetooth} \mbox{}\\

The proposed framework provides components of mruby bytecode loader using Bluetooth to use the function.
A component of mruby bytecode loader using Bluetooth is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages RiteVM configuration such as automatically generates the bytecode in the build description.
This generated bytecode is prepared beforehand in the SD card such as mruby libraries, and different from a bytecode transferred with Bluetooth.

@@

\end{lstlisting}
\caption{Main code for mruby bytecode loader using Bluetooth}
\label{maincode_mrubybluetooth}
\end{figure}
\subsection{Multitask}
\label{sec:Multitask}
This section describes implementation of multitasking in the proposed framework.
mruby on TECS has supported multitasking.
However, multitask processing in mruby on TECS requires the knowledge of the RTOS (TOPPERS/HRP2) for developers.

@@

In the proposed framework, as an approach for multitask processing, RiteVM scheduler is implemented.
The RiteVM scheduler is a fair scheduler and runs multiple tasks equally.

\subsubsection{RiteVM Scheduler} \mbox{}\\

The RiteVM scheduler is a cyclic handler, and calls {\it rotateReadyQueue}, a service call of $\mu$ITRON to switch tasks with the same priority.
In other words, the RiteVM scheduler calls {\it rotateReadyQueue} cyclically.
{\it rotateReadyQueue} is described as follows.

The case is assumed that there are two tasks with the same priority, and both tasks are in an infinite loop.
In the current system, when one task is executed first, another task would not be executed.
That is because the task with first execution runs in the loop.

@@

\subsubsection{Component of RiteVM Scheduler} \mbox{}\\

Figure \ref{fig:cyclic_handler} shows a component diagram of the cyclic handler.
The components of cyclic handler consist of two components: CyclicHandler and CyclicMain.
CyclicHandler {\it cell} configures the cyclic handler based on $\mu$ITRON.
Cyclic handlers based on $\mu$ITRON are described in detail \cite{par:microITRON}.
The cyclic handler has five arguments: ID, attribute, cyclic time, cyclic phase and access pattern.
CyclicHandler {\it cell} has these arguments as attributes of the {\it cell}.
CyclicMain {\it cell} is a component to perform the processing body of a cyclic handler.
{\it rotateReadyQueue} is implemented as the body.
Figure \ref{celltype_cyclic_handler} shows tCyclicMain {\it celltype}, which has a {\it call} port, an {\it entry} port and an attribute.
The {\it call} port is connected with the {\it entry} port of the Kernel {\it cell} ({\it tkernel.eiKernel}) to call functions of the kernel. 
The attribute is used as an arguments of {\it rotateReadyQueue}.

Figure \ref{build_cyclic_handler} shows a build description that corresponds to the component diagram shown in Figure \ref{fig:cyclic_handler}.
In the part of CyclicHandler {\it cell}, configurations of a cyclic handler is described such as attribute, cyclicTime and cyclicPhase.
In this case, the cyclic handler is executed when it is generated because the attribute is {\it TA\_STA} that represents the cyclic handler is in an operational state after the creation.
The cyclic handler is executed every one msec.
In another part, the priority of CyclicMain {\it cell} is described.
RITEVM\_PRIORITY defines the priority of mruby tasks.
In the main of CyclicMain, {\it rotateReadyQueue} is implemented and the priority is passed as the argument.

@@

\subsubsection{Synchronization of Multiple RiteVM Tasks} \mbox{}\\

In the proposed framework, RiteVMs read mruby bytecodes, and then execute the applications.
Eventflag, one of synchronous processing, is applied to synchronize the starting of multiple mruby applications.
Each task sets the flag pattern such as 0x01 (01) and 0x02 (10), and then waits the flag pattern, 0x3 (11), with AND.
This process can also apply to the case of the more tasks.
For example, in the case of the four RiteVM tasks, each task sets the flag pattern such as 0x01 (0001), 0x02 (0010), 0x04 (0100)  and 0x08 (1000), and then waits 0x0f (1111) with AND as shown in Figure \ref{fig:Eventflag} (A).

@@

\subsection{Benefits of Component-Based Development}
This section describes the design using the benefits of component-based development.

In the framework, RiteVM, RiteVM scheduler, and Eventflag are implemented as components.
Therefore, developers can easily add or remove these components and also reuse them.
For example, if developers do not need the cyclic handler, the .cdl files shown as Figures \ref{celltype_cyclic_handler} and \ref{build_cyclic_handler} should be commented out such as {\it //import($<$tCyclicHandler.cdl$>$);}.

In additon, the code size can desrease by developing with component-based. 
In the proposed framework, the advantage is applied in the Eventflag component.
The set pattern and wait pattern are defined as attributes of the component as shown in Figure \ref{fig:Eventflag} (B).
This design such as {\it cEventflag\_set(ATTR\_setPattern)} enables the program without ``if" statements and reuses the identical .c file.
Developers do not need to modify the .c file because the .cdl files are prepared in accordance with the number of RiteVMs
In additon, the components of Eventflag are built with {\it [optional]} in TECS.
{\it [optional]} means that the codes are run only when the call port is connected.
The .c file does not be rewritten even if developers do not use Eventflag.

\section{Experimental Evaluation}
\label{sec:Evaluation}
This section mentions experimental results and their consideration.
To analyze the advantages of the proposed framework, the evaluations are performed as follows.
\begin{itemize}
    \item Size and transfer time for mruby programs
    \item Execution time with singletasking, co-routine, and multitasking
    \item Overhead for cyclic time
    \item Code size using the benefit of CBD 
\end{itemize}

@@

\begin{table*}[t]
    \centering
    \caption{Comparion of the proposed and previous work}
    \begin{tabular}{c||c|ccccccc}
        & Bluetooth Loader & \shortstack{Call\\C function} & \shortstack{Legacy code of\\embedded system} & VM managenment & VM Scheduler & \shortstack{Synchronization of\\application} & Co-routine \\ \hline
        python-on-a-chip \cite{url:python-on-a-chip} &            &            &            &            &             &            & \checkmark \\
        Owl system \cite{par:owl}                    &            & \checkmark & Partially  &            &             &            & \checkmark \\
        eLua \cite{url:eLua}                         &            & \checkmark & Partially  &            &             &            & \checkmark \\
        mruby \cite{par:mruby}                       &            & \checkmark &            &            &             &            & \checkmark \\
        mruby on TECS \cite{par:mrubyonTECS}         &            & \checkmark & \checkmark & \checkmark & $\triangle$&            & \checkmark \\
        Proposed framework                           & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark & \checkmark \\
    \end{tabular}
    \label{tab:comparison}
\end{table*}
The open-source run-time systems for scripting languages have been proposed such as follow:
python-on-a-chip \cite{url:python-on-a-chip}, the Owl system \cite{par:owl}, eLua \cite{url:eLua}, mruby \cite{par:mruby}, \cite{url:mruby}, and mruby on TECS \cite{par:mrubyonTECS}.

@@

Table \ref{tab:comparison} shows a comparison between the proposed framework and previous work.

@@

The proposed framework is developed in component-base by TECS.
The facilities such as the RiteVM, the scheduler, and the Eventflag are implemented as components.
Developers can easily add or remove the facilities as necessary.
Component-based development can increase the productivity and decrease the complexity.
